---
title: "The National Agricultural Workers Survey (NAWS)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## NAWS data source
The source of this **NAWS** data: <https://www.doleta.gov/naws/pages/public-data/public-data-codebook-and-questionnaire.cfm>

```{r Load packages,include=FALSE}
library(plyr)
library(magrittr)
library(dplyr)
library(ggplot2)
library(caret)
library(lubridate)
```

## Load and combine data by farm workers ID
```{r Load and combine data by farm workers ID}
laborA<- data.frame(read.csv("~/Downloads/NAWS_A2E185.csv"))
laborB<- data.frame(read.csv("~/Downloads/NAWS_F2Y185.csv"))
intersect(names(laborA),names(laborB))
labor<-arrange( join(laborA,laborB) ,FWID)
laborA<-NULL
laborB<-NULL
labor<-subset(labor,select=-c(FWID,B14CODE,B15CODE,NT02COD,NQ02COD,PESTCONT))
labor<-labor[labor$FY==2015|2016,]
```

```{r Take a quick look at data,results='hide'}
dim(labor)
str(labor)
names(labor)
tail(labor,30)
```

### Preconditioning; convert the types of variables into factors
```{r,results='hide'}
nlev<-NULL
for (ii in 1:ncol(labor)) {
  nlev[ii]<-dim( table(labor[,ii]) )
  if ( nlev[ii] <=11 ) { 
    labor[,ii]<-factor( labor[,ii] )
  }
  else labor[,ii]<-as.numeric( labor[,ii] )
}
toFacNames<-c("Birth_State","B17","B18","B26a","B26b","B27a","B27b","G01","G02","A09","FY")
for (jj in 1:length(toFacNames)) {
  labor[,toFacNames[jj]]<-factor( labor[,toFacNames[jj]] )
}
```

### Plots about wage 
1. Wage as function of Age, 
2. grouped by CROP
```{r Make plots 1,warning=FALSE}
qplot(AGE,D12WG4, data=labor,geom=c("point","smooth"),main="Wage as function of Age") # smooth:loess
xyplot(D12WG4~AGE|CROP, data=labor,layout=c(6,1))
```

CROP: 0='NON-SASAGRICULTURE', 1='FIELD CROPS', 2='FRUITS&NUTS', 3='HORTICULTURE', 4='VEGETABLES', 5='MISC/MULT'

3. grouped by D37A: How far is your current job from your current residence?
( 1=I'm located at the job, 2=Within 9 miles, 3=10-24 miles, 
4=25-49 miles, 5=50-74 miles, 6=75 miles or more )

4. grouped by A21a: Does the farm worker have health insurance? ( 0=no, 1=yes, 7=don't know )

or A21b: Does the spouse have health insurance? ( 0=no, 1=yes, 7=don't know )
```{r Make plots 2,warning=FALSE}

xyplot(D12WG4~AGE|D37A, data=labor,layout=c(6,1))
#qplot(AGE,D12WG4, data=labor,color=D37A)

xyplot(D12WG4~AGE|A21a, data=labor,layout=c(6,1))

qplot(D12WG4,data=labor,geom="density",color=A21a)
```

Check whether the farm worker and their spouses have health insurance:
A21a: Does the farm worker have health insurance? ( 0=no, 1=yes, 7=don't know )
A21b: Does the spouse have health insurance? ( 0=no, 1=yes, 7=don't know )
```{r}
qplot(A21a,data=labor,fill=A21b)
```

## Hypothesis testing for paired categorical variables
NP1F: In the last 12 months, have you loaded, mixed or applied pesticides?
```{r Hypothesis testing, warning=FALSE}
co<-NULL
labor$NP01f<-as.factor(labor$NP01f)

for (ii in 1:ncol(labor)) {
  laborT<-as.data.frame(labor$NP01f)
  
  if ( dim(table(labor[,ii]))<=11 ) { 
  
    laborT[,2]<-labor[,ii]
    laborT[,2]<-as.factor( laborT[,2] )
    names(laborT)[2]<-names(labor)[ii]
    #print(names(labor)[ii])
    bad<-rowSums( is.na(laborT) | laborT=="" )>0
    laborT<-laborT[!bad,]
    
    chisqRes<-chisq.test(table(laborT))
    co[ii]<-chisqRes$p.value
    #if (co[ii]>0.05){print(names(labor)[ii])}
  }
}
plot(co)
print(names(labor)[!is.na(co) & co<0.05])
#plot(co,ylim=c(0,1))
```

## Training with Random Forest
Here we select group B variables (Adult Education, Languages, Work Experience, Residence) to make models, predictions, confusion matrices

```{r Training with Random Forest, group B variables: Adult Education, Languages, Work Experience, Residence}
labor %>% filter(FY==2016) %>% select( grep("^B0",names(labor)),NP01f) ->laborC

for (ii in 1:ncol(laborC)) {
  bad<-rowSums( is.na(laborC) | laborC=="" )>0
  laborC<-laborC[!bad,]
}

inTrain <- createDataPartition(y=laborC$NP01f, p=0.7, list=FALSE)
training <- laborC[inTrain, ]
testing <- laborC[-inTrain, ]
dim(training)
dim(testing)

table(training$NP01f,useNA="ifany")

ptm <- proc.time()
modRF<-train(NP01f~.,data=training,method="rf",ntree=10,tuneGrid=data.frame(mtry=6))
proc.time()-ptm

predRF<-predict(modRF,testing)
confusionMatrix(predRF,testing$NP01f)
plot(modRF$finalModel$importance)
modRF$finalModel$importance
```
B17: [IF FOREIGN BORN:] In what country (outside of the U.S.A.) did you live before coming to the U.S.A?

Here we select group G variables (Income and Assets) to make models, predictions, confusion matrices.

```{r Training with Random Forest, group G variables: Income and Assets}
labor %>% filter(FY==2016) %>% select( grep("^G0",names(labor)),NP01f) ->laborC

for (ii in 1:ncol(laborC)) {
  bad<-rowSums( is.na(laborC) | laborC=="" )>0
  laborC<-laborC[!bad,]
}

inTrain <- createDataPartition(y=laborC$NP01f, p=0.7, list=FALSE)
training <- laborC[inTrain, ]
testing <- laborC[-inTrain, ]
dim(training)
dim(testing)

table(training$NP01f,useNA="ifany")

ptm <- proc.time()
modRF<-train(NP01f~.,data=training,method="rf",ntree=10,tuneGrid=data.frame(mtry=6))
proc.time()-ptm

predRF<-predict(modRF,testing)
confusionMatrix(predRF,testing$NP01f)
plot(modRF$finalModel$importance)
modRF$finalModel$importance
```

G3: What was your FAMILY'S TOTAL INCOME last year in U.S.A. dollars (U.S.A. earners only)?

```{r}

```

